---
title: "Analysing predictions"
author: "Mateusz Żółtak"
date: "4 maja 2017"
output: html_document
---

Raw data:
```{r}
metgisRaw = tbl(src, sql("SELECT * FROM swb.metgis_data"))
head(metgisRaw)
```

Forecasts are provided for 3h periods.

Take data for point_id 1, limit to predictions yunger then 3 days before the actual 3h period, then for each 3h period compute maximum difference between the actual measure and predictions.
Finally plot histogram of this differences.

```{r}
tmp = metgisRaw %>%
  filter(point_id == 1) %>%
  collect() %>%
  group_by(point_id, date) %>%
  arrange(desc(download_date)) %>%
  filter(download_date - first(download_date) >= -60*60*24*4) %>%
  summarize(n = n(), diffMax = max(abs(first(tmp2m) - tmp2m))) %>%
  arrange(date)
tmp %>%
  ggplot(aes(x = diffMax)) +
  geom_density()
```

Take data for point_id 1, limit to predictions yunger then 3 days before the actual 3h period, then for each prediction for each 3h period compute its difference with the actual measure.

```{r}
tmp2 = metgisRaw %>%
  filter(point_id == 1) %>%
  collect() %>%
  group_by(point_id, date) %>%
  arrange(desc(download_date)) %>%
  filter(download_date - first(download_date) >= -60*60*24*3) %>%
  mutate(order = -(1:n()) + 1, diff = first(tmp2m) - tmp2m) %>%
  filter(download_date != first(download_date), as.character(date) < '2017-05-04')
tmp2 %>%
  arrange(desc(date), desc(download_date))
```

Take data for point_id 1, aggregate 3h periods to full days by taking average, limit to predictions yunger then 3 days before the actual date, then for each prediction for each date compute its difference with the actual measure.

```{r}
tmp3 = metgisRaw %>%
  filter(point_id == 1) %>%
  collect() %>%
  mutate(date = substr(date, 1, 10)) %>%
  group_by(point_id, download_date, date) %>%
  summarize(tmp2m = mean(tmp2m))
tmp3 = tmp3 %>%
  group_by(point_id, date) %>%
  arrange(desc(download_date)) %>%
  filter(download_date - first(download_date) >= -60*60*24*3) %>%
  mutate(diff = round(first(tmp2m) - tmp2m, 3)) %>%
  filter(download_date != first(download_date), as.character(date) < '2017-05-04')
tmp3 %>%
  arrange(desc(date), desc(download_date))
```
