---
title: "First analysis"
author: "Susanna Galimberti"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# import libraries
library(dplyr)
library(ggplot2)

# import our helper functions
source('functions.R')

# connect to the database
source('.password.R')
src = src_postgres("eo4water", "ivfl-arc.boku.ac.at", user = "weather", password = pswd)
```

Our point of interest is (16.56972 E, 48.11028 N) and we are interested in the period 2016-09-01 - 2016-10-31

```{r gettingData}
pointX = 16.56972
pointY = 48.11028
dateFrom = '2016-09-01'
dateTo = '2016-10-31'

dataWide = get_data_wide(src, pointX, pointY, dateFrom, dateTo)
dataWide
dataLong = get_data_long(src, pointX, pointY, dateFrom, dateTo)
dataLong

# and old-style
bokuData = get_meteo_data(src, pointX, pointY, 'boku') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
bolamData = get_meteo_data(src, pointX, pointY, 'bolam') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
metgisData = get_meteo_data(src, pointX, pointY, 'metgis') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
openweatherData = get_meteo_data(src, pointX, pointY, 'openweather') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
```

# TEMPERATURE PLOT

Old R graphics style:

```{r}
plot(x = bokuData$date, y = bokuData$temp, type = "l", col = "green")
lines(x = bolamData$date, y = bolamData$temp, type = "l", col = "red")
lines(x = metgisData$date, y = metgisData$temp, type = "l", col = "blue")
lines(x = openweatherData$date, y = openweatherData$temp, type = "l", col = "orange")
# how to show the whole plot?
```

Using ggplot:

```{r}
dataLong %>%
  ggplot(aes(x = date, y = temp, group = source, color = source)) +
    geom_line()
```

# RAIN PLOT

Old R graphics style:

```{r}
plot(x = bokuData$date, y = bokuData$rain, type = "p", col = "green")
lines(x = bolamData$date, y = bolamData$rain, type = "p", col = "red")
lines(x = metgisData$date, y = metgisData$rain, type = "p", col = "blue")
lines(x = openweatherData$date, y = openweatherData$rain, type = "p", col = "orange")
# how to plot just rain value > 0 ?
```

Using ggplot:

```{r}
dataLong %>%
  filter(rain > 0) %>%
  ggplot(aes(x = date, y = rain, group = source, color = source)) +
    geom_point()
```

# TEMPERATURE DESCRIPTIVE STATISTICS

```{r, echo = TRUE}
summary(dataWide$bk_temp)
summary(dataWide$bl_temp)
summary(dataWide$ow_temp)
summary(dataWide$mg_temp)
```

## testing normal distribution

```{r, echo=TRUE}
shapiro.test(dataWide$bk_temp)
```

Data not normally distributed (we reject NULL HYP.) -> non-parametric test

```{r, echo=TRUE}
wilcox.test(dataWide$bk_temp, dataWide$bl_temp, paired=T)
wilcox.test(dataWide$bk_temp, dataWide$mg_temp, paired=T)
wilcox.test(dataWide$bk_temp, dataWide$ow_temp, paired=T)
```

IF (!!) this is correct, then it tells that only bl_temp made good prognoses

```{r, echo=TRUE}
tempByCol$diff_bktemp <- c(tempByCol$bk_temp - tempByCol$bl_temp)
```

difference Tboku e T bolam

```{r, echo=TRUE}
shapiro.test(tempByCol$diff_bktemp)
```

normally distributed -> t-test

```{r, echo=TRUE}
t.test(tempByCol$diff_bktemp, mu=0)
```

null hyp. = the mean is equal to 0 and we cannot reject it --> random
