---
title: "First analysis"
author: "Susanna Galimberti"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# import libraries
library(dplyr)
library(ggplot2)

# import our helper functions
source('functions.R')

# connect to the database
source('.password.R')
src = src_postgres("eo4water", "ivfl-arc.boku.ac.at", user = "weather", password = pswd)
```

Our point of interest is (16.56972 E, 48.11028 N) and we are interested in the period 2016-09-01 - 2016-10-14

```{r gettingData}
pointX = 16.56972
pointY = 48.11028
dateFrom = '2016-09-01'
dateTo = '2016-10-14'

dataWide = get_data_wide(src, pointX, pointY, dateFrom, dateTo)
dataWide # organize data in colomns (bk_temp, bl_temp,...)
dataLong = get_data_long(src, pointX, pointY, dateFrom, dateTo)
dataLong # organize data in rows (source)

# and old-style
bokuData = get_meteo_data(src, pointX, pointY, 'boku') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
bolamData = get_meteo_data(src, pointX, pointY, 'bolam') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
metgisData = get_meteo_data(src, pointX, pointY, 'metgis') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
openweatherData = get_meteo_data(src, pointX, pointY, 'openweather') %>%
  filter(date >= dateFrom & date <= dateTo) %>%
  arrange(date)
```

# TEMPERATURE PLOT

Old R graphics style:

```{r}
plot(x = bokuData$date, y = bokuData$temp, type = "l", col = "green")
lines(x = bolamData$date, y = bolamData$temp, type = "l", col = "red")
lines(x = metgisData$date, y = metgisData$temp, type = "l", col = "blue")
lines(x = openweatherData$date, y = openweatherData$temp, type = "l", col = "orange")
# how to show the whole plot?
```

Using ggplot:

```{r}
dataLong %>%
  ggplot(aes(x = date, y = temp, group = source, color = source)) +
  geom_line() +
  scale_x_date(date_breaks = "10 days") 
```
# PAIRWISE DIFFERENCES OF AVERAGE DAILY TEMPERATURE
dif: We create the new coloumn "dbkbl_temp"" to get the pairwise differences and we plot them.
  "geom_line" connect data in order of the x axsis.
dif2: We create new coloumns  
  "lead" or "lag" compare value offset by previous or next value.
  "group"or "ungroup" only difference is that appear the number of groups: date [44], we group at the beggining to compare and after   we retidy up the rows per      source.
```{r}
dif = dataWide %>%
  mutate(dbkbl_temp = bk_temp - bl_temp) 

dif %>%
  ggplot(aes(x = date, y = dbkbl_temp)) +
  geom_line() 

dif2 = dataLong %>%
  group_by(date) %>%
  arrange(source) %>%
  mutate(
    diff_temp = temp - lead(temp),
    diff_label = paste(source, lead(source))
  ) %>%
  ungroup() 

dif2 %>%
  filter(!is.na(diff_temp)) %>%
  ggplot(aes(x = date, y = diff_temp, group = diff_label, color = diff_label)) +
  geom_line()

dif2 %>%
  filter(!is.na(diff_temp)) %>%
  ggplot(aes(x = diff_temp, group = diff_label, color = diff_label)) +
  geom_density()

```

# DESCRIPTIVE STATISTICS
```{r}
dif2 %>%
  filter(diff_label == 'bk bl') %>%
  select(diff_temp) %>%
  summary()
dif2 %>%
  filter(diff_label == 'bl mg') %>%
  select(diff_temp) %>%
  summary()
dif2 %>%
  filter(diff_label == 'mg ow') %>%
  select(diff_temp) %>%
  summary()


dif2 %>%
  filter(!is.na(diff_temp)) %>%
  group_by(diff_label) %>%
  ggplot(aes(x = diff_label, y = diff_temp)) + 
  geom_boxplot()+
  scale_y_continuous(breaks = c(-4, -3.5, -3, -2.5, -2, -1.5, -1, -0.5, -0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4))

```


# RAIN PLOT

Old R graphics style:

```{r}
plot(x = bokuData$date, y = bokuData$rain, type = "p", col = "green")
lines(x = bolamData$date, y = bolamData$rain, type = "p", col = "red")
lines(x = metgisData$date, y = metgisData$rain, type = "p", col = "blue")
lines(x = openweatherData$date, y = openweatherData$rain, type = "p", col = "orange")
```

Using ggplot:

```{r}
dataLong %>%
  filter(rain > 0) %>%
  ggplot(aes(x = date, y = rain, group = source, color = source)) +
    geom_point() +
  scale_x_date(date_breaks = "10 days")
```

# PAIRWISE DIFFERENCES OF SUM DAILY PRECIPITATION
```{r}
difRAIN = dataLong %>%
  group_by(date) %>%
  arrange(source) %>%
  mutate(
    diff_rain = rain - lead(rain),
    diff_rain_label = paste(source, lead(source))
  ) %>%
  ungroup() 

difRAIN %>%
  group_by(diff_rain_label) %>%
  count(
    diff_rain > 0,
    diff_rain < 0,
    diff_rain = 0
    ) # attempt 1 = useless!

difRAIN %>%
  filter(!is.na(diff_rain)) %>%
  ggplot(aes(x = date, y = diff_rain, group = diff_rain_label, color = diff_rain_label, fill = diff_rain_label)) + 
  geom_col(position = "dodge")
```

bl compared to bk overestimates the amount of daily precipitation.
mg compared to bl underestimates the amount of daily rain --> compare mg-b!
ow compared to mg overestimates --> trend similar to bl?


```{r}
difRAIN %>%
  filter(!is.na(diff_rain)) %>%
  group_by(diff_rain_label)%>%
  ggplot(aes(x = diff_rain, group = diff_rain_label, color = diff_rain_label)) +
  geom_density() # not good

difRAIN %>%
  filter(!is.na(diff_rain)) %>%
  group_by(diff_rain_label) %>%
  ggplot(aes(x = diff_rain_label, y = diff_rain)) + 
  geom_boxplot() # how to show it better?


```
# SAME ANALYSIS FOR NEUSIEDLER SEE POINT
The poin of interest now is (16.745 E, 47.929 N) and we are interested in the period 2016-09-01 - 2016-10-14.
This point is settled on the shore of a big lake.

```{r gettingData}
pointX = 16.745
pointY = 47.929
dateFrom = '2016-09-01'
dateTo = '2016-10-14'

dataWideNS = get_data_wide(src, pointX, pointY, dateFrom, dateTo)
dataWideNS # organize data in colomns (bk_temp, bl_temp,...)
dataLongNS = get_data_long(src, pointX, pointY, dateFrom, dateTo)
dataLongNS # organize data in rows (source)
```

# TEMPERATURE NS
```{r}
dataLongNS %>%
  ggplot(aes(x = date, y = temp, group = source, color = source)) +
  geom_line() +
  scale_x_date(date_breaks = "10 days") 
```

# PAIRWISE DIFFERENCES NS
```{r}
difNS = dataWideNS %>%
  mutate(dbkbl_tempNS = bk_temp - bl_temp) 

difNS %>%
  ggplot(aes(x = date, y = dbkbl_tempNS)) +
  geom_line() 

dif2NS = dataLongNS %>%
  group_by(date) %>%
  arrange(source) %>%
  mutate(
    diff_tempNS = temp - lead(temp),
    diff_labelNS = paste(source, lead(source))
  ) %>%
  ungroup() 

dif2NS %>%
  filter(!is.na(diff_tempNS)) %>%
  ggplot(aes(x = date, y = diff_tempNS, group = diff_labelNS, color = diff_labelNS)) +
  geom_line()

dif2NS %>%
  filter(!is.na(diff_tempNS)) %>%
  ggplot(aes(x = diff_tempNS, group = diff_labelNS, color = diff_labelNS)) +
  geom_density()

```
 
# DESCRIPTIVE STATISTICS NS
```{r}
dif2NS %>%
  filter(diff_labelNS == 'bk bl') %>%
  select(diff_tempNS) %>%
  summary()
dif2NS %>%
  filter(diff_labelNS == 'bl mg') %>%
  select(diff_tempNS) %>%
  summary()
dif2NS %>%
  filter(diff_labelNS == 'mg ow') %>%
  select(diff_tempNS) %>%
  summary()


dif2NS %>%
  filter(!is.na(diff_tempNS)) %>%
  group_by(diff_labelNS) %>%
  ggplot(aes(x = diff_labelNS, y = diff_tempNS)) + 
  geom_boxplot()+
  scale_y_continuous(breaks = c(-4, -3.5, -3, -2.5, -2, -1.5, -1, -0.5, -0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4))

```

